Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/Irhe}{https://github.com/karhunenloeve/Irhe}.}

### 2020-06-17 llncs2e/paper12.tex
corrections

\section{Related work}
The related work is divided into two sections. In the first section we present chronologically the milestones in the consideration of neural nets from a differential geometric perspective. In the second section we present work on earlier attempts to determine their minimum embedding dimension.

\subsubsection*{Riemannian geometry in neural nets} The manifold of a dense neural net is in most cases Euclidean. Very recently, there have been results in the direction of other manifolds on which neural nets can operate. The spherical neural nets operate on a manifold with commutative group structure \cite{CohenGKW18,EstevesAMD20}.

A metric is required to describe the geometric properties of such a manifold. The change of coordinate systems and the learned metric tensor of a Riemann manifold during back propagation of neural nets has been investigated \cite{HauserR17}. Further, it has been shown that dense neural nets can't approximate all functions arbitrarily precisely \cite{Johnson19}. In response to this result, neural nets with residual connections have been examined. Residual nets add the output of one layer to a deeper one and therefore bypass some of the layers in between. This type of neural net is indeed an universal function approximator \cite{LinJ18}.

\subsubsection*{Embedding dimension of neural nets} Cybenko showed that dense neural nets with a single hidden layer can approximate any continuous function with a bounded domain with arbitrary small error \cite{Cybenko92}. The upper bound of the embedding dimension for ReLU nets has recently been proven to be $(n+4)$ with $n$ being the sample dimension \cite{RaghuPKGS17}. Bartlett et al. show that a nearly-tight bound for the VC dimension of a neural net with $W$ weights and $L$ layers has $O(WL \log W)$ but $\Omega(WL \log W/L)$ VC dimensions \cite{BartlettHLM19}.

The depth of a neural net has been determined to be a parameter depending on the moduli of continuity of a function to be approximated \cite{LinJ18}. Inspired by the theory of finite difference methods, such a residual layer has been defined as a forward or backward difference operator on a partition of the layers. The state space dimension with residual connections is homeomorphic to $\mathbb{R}^{k \cdot n}$, where $n = \text{dim} \; M$ and $k$ is the number of times the difference operator was used \cite{HauserGJR19}. Thus, such a layer is capable of embedding into $k\cdot n$ dimensional space. The view that a neural net operates on a smooth manifold inspired us to investigate the invariants of this manifold for the theory discussed in this paper.

To these fundamental results we contribute a topological approach to parameterization problems. Similar to Futagami et al. \cite{FutagamiYS19}, we study the manifold on which we suspect the data. Our approach is applicable to topological spaces in general and is based on persistent homology. We are looking for the dimension of a space with the same topological properties as data indicates and of which we know that it can be approximated. We use the simplest decomposition into a product space of realaxes and $1$-spheres to obtain the same homology groups as the filtration of the data has. Thus, we obtain a computable lower bound for the embedding dimension of any neural net with differentiable structure.

\newpage
\section{Smooth manifolds, Lie groups and persistent homology}
The assumption that a set of points lies on a manifold is more accurate than handling points in Euclidean space. A \emph{topological manifold} $M$ is a \emph{Hausdorff} space, which means that at any two points $x,y \in M$ there are always two open neighborhoods $U_x,U_y \subset M$, so that their intersection is empty. This behaviour provides the intuitive feeling of the position of points in space. The topology $(M, \tau)$ is a set system that describes the structure of a geometric object. Thereby $M$ itself and $\emptyset$ must be contained in $\tau$, and any arbitrary union of open sets from $\tau$ and any intersections from $\tau$ must be in $\tau$. If this set system has a smaller set that generates the topology $\tau$ with the aid of finitely many unions, then this smaller set is a \emph{basis} of the topology. We demand of a manifold that the basis is at most countable. At last, $M$ is locally Euclidean, i.e. each point has a neighborhood which can be homeomorphically mapped to a subset of $\mathbb{R}^n$. The integer $n$ is the dimension of $M$. Further, we demand some additional smooth structure.

\subsection{Smooth manifolds} A smooth structure is a stronger condition than a topological manifold which can be described by a family of continuous functions. This assumption can be justified by the theorem of Stone-Weierstrass which proves that any continuous function can be approximated arbitrarily precisely by a smooth one \cite{stone1948generalized}. Smooth manifolds are described by a collection of local coordinate charts $\varphi: U \rightarrow \varphi(U) \subseteq \mathbb{R}^n$, homeomorphic to $\mathbb{R}^{n}$ and an open neighborhood $(U_i,\varphi_i)$ in $M$ which is covered by this map. The family of such maps that cover $M$ is the \textit{atlas} $\mathscr{A} = (U_i,\varphi_i)_{i \in I}$ \cite{lee2013smooth}. The functions $x_1, \cdots, x_n: U \rightarrow \mathbb{R}$ are called local coordinates $\varphi(p) = (x_1(p), \cdots, x_n(p))$. If the atlas is maximal with respect to inclusion, then it is a \emph{differentiable structure}. Each atlas for $M$ is included in a maximal atlas. Because of the differentiable structure, the maps in $\mathscr{A}$ are also compatible with all other maps of $\mathscr{A}_{\text{max}}$. Thus, it is sufficient to describe smooth manifolds by an atlas that is not maximal.

As long as the activation functions are smooth and of the same differentiable structure, they are also compatible and the manifold does not change while propagating the data trough the layers. Only the coordinate representation changes.

@@

We apply a theorem from Lie theory, which states that each connected Abelian Lie group $(M,\bullet)$ of dimension $\dim M = n$ is isomorphic to a product space $\mathbb{R}^p \times \mathbb{T}^q$ with $p+q =n$, for a proof we refer to \cite[p.~116]{onishchik1993lie}. The $q$-torus $\mathbb{T}^q$ is a surface of revolution, which means it moves a curve around an axis of rotation. These axes are given by $1$-spheres, such that the $q$-torus is a product space $\mathbb{T}^q = \mathcal{S}^1_1 \times \cdots \times \mathcal{S}^1_q$. The initial decomposition of connected commutative Lie groups can be further simplified into $M \simeq \mathbb{R}^p \times (\mathcal{S}^1_1 \times \cdots \times \mathcal{S}^1_q)$. Recall, that the $(n-1)$-sphere is given by $\mathcal{S}^{n-1} := \{x \in \mathbb{R}^n \; \vert \; \norm{x}_2 = 1\}$. Thus, $\mathcal{S}^1$ can be embedded into a two-dimensional Euclidean space. Under our assumption, we can derive how many dimensions are at least needed for a suitable embedding. For each $1$-sphere we count two dimensions and for each real line accordingly one. Finally, we have to estimate from data how many factor spaces are necessary to represent the topology of the underlying data manifold sufficiently accurately.

\subsection{Persistent homology}
Algebraic topology provides a computable tool to study not the topology of a set of points itself but an Abelian group which is attached to the topology. The core interest in this discipline lies in \emph{homotopy equivalences} of an equivalence class to which objects belong that are continuously deformable into one another. This term is much broader than a homeomorphism. For two topological spaces $X$ and $Y$ we are looking for a function $h: X \times I \rightarrow Y$, which gives the identity at the time $h_0(X) = X$ and for $h_1(X) = X$ a mapping into the other topological space. If this mapping is continuous with respect to its arguments, it is called \emph{homotopy}. Let us consider two functions $f,g: I \rightarrow X$ so that $f(1) = g(0)$. Then there is a \emph{composition} of product paths $f \cdot g$, which pass first through $f$ and then through $g$ and which is defined as $f \cdot g (s) = f(2s)$ for $0 \leq s \leq 1/2$ and $g(2s - 1)$, for $1/2 \leq s \leq 1$. We first run $f$ at double speed up to $1/2$ and from $1/2$ to $1$ we run the function $g$ at double speed. In addition, suppose that we have a family of functions $f: I \rightarrow X$ that have the same start and end point $f(0) = f(1) = x_0 \in X$. These functions intuitively form a \emph{loop}. The set of all homotopy classes $[f]$ of loops $f: I \rightarrow X$ at base point $x_0$ is noted as $\pi_1(X,x_0)$ and is called first homotopy group or fundamental group. The group operation is the product $[f][g] = [f \cdot g]$ of equivalence classes. If we do not look at the interval $I$, but at mappings considering the unit cube $I^n$, we obtain the $n$-th homotopy group $\pi_n(X,x_0)$ by analogy. With the help of the homotopy groups we can study the connected components of a topological space for the $0$th group, the loops for the $1$th group, the cavities for the $2$th and so forth. Since homotopy groups are very difficult to compute, we resort to an algebraic variant, the \emph{homology groups}.

@@

\input{figure1.tex}
\subsubsection{Simplicial complexes} A \emph{simplicial complex} $K$ is a set of simplices, such that any face of a simplex of $K$ is a simplex of $K$ and the intersection of any two simplices of $K$ is either empty or a common face of both \cite[p.~11]{boissonnat2018geometric}. Further, we require every $(k+1)$-element, with $k < n$ to be a $k$-simplex of $[\sigma] \subset K$. This definition means that the simplicial complex is also a topological space. The properties described ensure that simplices are added to this set in a certain way, namely so that they are only \emph{glued} edge to edge, point to point, and surface to surface. Thus, every $n$-simplex has $n+1$ distinct vertices and no other $n$-simplex has the same set of vertices.  This is a unique combinatorial description of vertices together with a collection of sets $\{[\sigma]_1, \cdots, [\sigma]_m\}$ of $k$-simplices, which are $k+1$ element subsets of $K$ \cite[p.~107]{hatcher2002algebraic}. This set system can be realized geometrically with a homeomorphism, by a corresponding mapping into an Euclidean space.

\subsubsection{Associated Abelian groups} Algebraically, the simplices can be realized as a system of linear combinations as $\sum_i \lambda_i [\sigma]_{i}$ following Eq. \ref{simplex}. Together with addition as an operation, they form a group structure called $k$th chain group $\langle C_k, +\rangle$ for the corresponding dimension of the $k$-simplices used. The group is commutative or Abelian. The used coefficients $\lambda_i$ induce an orientation, intuitively a direction of the edges is given here by negative and positive signs, so that $[\sigma] = -[\tau]$, iff $[\sigma] = [\tau]$ and $[\sigma]$ and $[\tau]$ have different orientations. The objects of the chain group are called $k$-chains, for $[\sigma] \in C_k$. The groups are connected by a homomorphism, a mapping which respects the algebraic structure, the boundary operator: $\partial_k: C_k \rightarrow C_{k-1}$, $\partial_k [\sigma] \mapsto \sum_i (-1)^i [v_0,v_1,\cdots,\hat{v}_i, \cdots, v_n]$. The $i$th face is omitted, alternatingly. Using the boundary operator we yield the following sequence of Abelian groups, which is a chain complex cf. Fig. \ref{figure1} a):
\begin{equation}
    0 \xrightarrow[]{\partial_{k+1}} C_k \xrightarrow[]{\partial_{k}} C_{k-1} \xrightarrow[]{\partial_{k-1}} \cdots \xrightarrow[]{\partial_{2}} C_1 \xrightarrow[]{\partial_{1}} C_0 \xrightarrow[]{\partial_{0}} 0.
\end{equation}

@@

\section{Neural nets}
\label{sec:neuralnetworks}
Neural nets are a composition of \emph{affine transformations} with a non-linear activation function. This transformation obtains collinearity. It also preserves parallelism and partial relationships. The projection onto the $(l+1)$-th layer of such a network can be formalized as $\textbf{x}^{(l+1)} = f(\mathbf{W}^{(l+1)} \cdot \textbf{x}^{(l)}+\textbf{b}^{(l+1)}).$ The composition of multiple such maps is a deep neural net.

@@

\section{Experimental setting}
We train an auto-encoder using as input heavily noisy images and map them to their noiseless original. We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$, such that the input is weighted $0.5 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$. The embedding dimension is increased for each experiment by $2$, i.e. $2,4,6, \cdots, 784$. $392$ neural nets have been trained $\times 100$. Each line of Fig. \ref{figure2} c), d) represents one neural architecture as averaged loss function.
\input{figure2.tex}
\subsubsection*{Persistence landscapes} We use the Delaunay-complex for the filtration, according to Melodia et al. \cite{lume}. The maximum $\alpha$-square is set to $e^{12}$. The maximal expansion dimension of the simplicial complex is set to $10$. The maximal edge length is set to $0.1$. The persistent landscapes are smoothened by a Gaussian filter $G(x) = 1 / \sqrt{2 \pi \sigma^2} \cdot e^{- x^2 / 2 \sigma^2}$ with $\sigma = 2$ for visualization purposes. We implement the persistence diagrams and persistence landscapes using the \texttt{GUDHI} library \cite{gudhi:urm}. Persistent homology is computed on a \texttt{NVIDIA Quadro P4000}.
\subsubsection*{Hyperparameters} The batch size is set to $128$, which should help to better recognize good parametrization. Higher batch size causes explosions of the gradient, which we visualize in Fig. \ref{figure2} c), d). For optimization we use \texttt{ADAM} \cite{KingmaB14} and a learning rate of $10^{-4}$. The dataset is randomly shuffled. We implement in \texttt{Python} $3.7$. We use \texttt{Tensorflow} as backend \cite{tensorflow2015-whitepaper} with \texttt{Keras} as wrapper \cite{chollet2015keras}. The neural training is conducted on \texttt{Intel Core i}$7$ $9700$\texttt{K} processors.
\input{figure3.tex}
\subsubsection*{Neural net} We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = 1/n \sum_{i=1}^{n} (\textbf{x}^{(0)}_i-\textbf{y}_i)^2$ as loss function. The sigmoid activation function $\sigma(\mathbf{x}^{(l)}) = 1 / 1 + e^{-\mathbf{x}^{(l)}}$ is applied throughout the net to yield the structure of a smooth manifold. Dense neural nets are used with bias term. The datasets are flattened from $(32,32,3)$ into $(3072)$. The first layer is applied to the flattened dataset, which is also adapted by back propagation, i.e. $\varphi^{(0)}(x^{(0)}) =$ $\sigma(\sum_{i=1}^{n} \textbf{W}_{ji}^{(1)} \textbf{x}_{i}^{(0)} + \textbf{b}_{i}^{(1)})$. It transfers the data to the desired embedding dimension. The subsequent layers are implemented as residual invertible layers following Dinh et al. \cite{DinhSB17}. We use invertible architectures to avoid degenerated cases and to stay on the same differentiable structure during learning within the layers to be tested. A total of $5$ of these hidden layers are used, all with the same embedding dimension. This diagram describes our architecture:
    \begin{align*}
    x^{(0)}(M) \xrightarrow[]{\varphi^{(0)}}
    x^{(1)}(U) \xleftrightarrow[]{\varphi^{(1)}}
    x^{(2)}(U) \xdashleftrightarrow[]{\varphi^{(L-1)}} 
    x^{(L-1)}(U) \xhookrightarrow[]{\varphi^{(L)}} 
    x^{(L)}(M).
    \end{align*}

>>>>

\section{Related work}
We present chronologically relevant work on neural nets from a differential geometric perspective. Further, we summarize earlier attempts to determine depth and width of a neural net, depending on the data to be processed.

\subsubsection*{Differential geometry in neural nets} The manifold of a dense neural net is in most cases Euclidean. Recently there have been results in the direction of other manifolds on which neural nets can operate. The spherical neural nets operate on a manifold with a commutative group structure \cite{CohenGKW18,EstevesAMD20}.

A metric describes the geometric properties of a manifold. The change of coordinate systems and the learned metric tensor of a smooth manifold during backpropagation was formalized \cite{HauserR17}. Further, it has been shown that dense neural nets cannot approximate all functions arbitrarily precisely \cite{Johnson19}. Thereupon neural nets with residual connections were investigated. Residual nets add the output of one layer to a deeper one, bypassing some of the layers in between. Indeed, they are universal function approximators \cite{LinJ18}. Inspired by finite difference methods, such a residual layer was defined as a forward or backward difference operator on a partition of layers. The state space dimension with residual connections is homeomorphic to $\mathbb{R}^{k \cdot n}$, where $n = \text{dim} \; M$ and $k$ is the number of times the difference operator was used \cite{HauserGJR19}. Such a layer is able to embed in $k\cdot n$ dimensional Euclidean space. The smooth manifold perspective inspired us to investigate its invariants leading into the theory discussed in this paper.

\subsubsection*{Embedding dimension of neural nets} Cybenko showed that dense neural nets with one hidden layer can approximate any continuous function with a bounded domain with arbitrarily small error \cite{Cybenko92}. The upper bound of the embedding dimension for ReLU nets was recently determined to be $(n+4)$, where $n$ is the sample dimension \cite{RaghuPKGS17}. Bartlett et al. show an almost tight bound for VC dimension. A neural net with $W$ weights and $L$ layers has $O(WL \log W)$ but $\Omega(WL \log W/L)$ VC dimensions \cite{BartlettHLM19}. Its depth was determined as a parameter depending on the moduli of continuity of the function to be approximated \cite{LinJ18}.

To these fundamental results we contribute a topological approach to parameterization. Similar to Futagami et al.\cite{FutagamiYS19}, we investigate the manifold on which we suspect the data. Our approach is applicable to topological spaces in general and is based on persistent homology. We look for the dimension of a space with the same topological properties that the data indicates and which can be approximated. We assume a simple decomposition into a product space of realaxes and $1$-spheres to get the same homology groups as the persistent ones on a filtration of data. The dimension of the decomposition is a computable lower bound.

\section{Smooth manifolds, Lie groups and persistent homology}
The assumption that a set of points lies on a manifold is more accurate than the treatment in Euclidean space. A \emph{topological manifold} $M$ is a \emph{Hausdorff} space, which means that at any two points $x,y \in M$ have always open neighborhoods $U_x,U_y \subset M$, so that their intersection is empty. This behaviour gives an intuitive feeling for the position of points in space. The topology $(M, \tau)$ is a set system that describes the structure of a geometric object. Here $M$ itself and $\emptyset$ must be contained in $\tau$, and any union of open sets from $\tau$ and any intersection of $\tau$ must be contained in $\tau$. If this set system has a subset which generates the topology $\tau$ with finitely many unions, then this subset is a \emph{basis} of the topology. We demand of a manifold that the basis is at most countable. In addition, $M$ is locally Euclidean, i.e. each point has a neighborhood which can be mapped homeomorphically to a subset of $\mathbb{R}^n$. The integer $n$ is the dimension of $M$.

\subsection{Smooth manifolds}
A smooth structure is a stronger condition than a topological manifold which can be described by a family of continuous functions. This assumption is justified by the theorem of Stone-Weierstrass which proves that every continuous function can be approximated arbitrarily exactly by a smooth one \cite{stone1948generalized}. Smooth manifolds are described by a family of local coordinate maps $\varphi: U \rightarrow \varphi(U) \subseteq \mathbb{R}^n$ which are homeomorphic to a subset of $\mathbb{R}^{n}$ and cover an open neighborhood $(U_i,\varphi_i)$ in $M$. This family of such charts is called \emph{atlas} on $M$, i.e. $\mathscr{A} = (U_i,\varphi_i)_{i \in I}$ \cite{lee2013smooth}. The functions $x_1, \cdots, x_n: U \rightarrow \mathbb{R}$ are called local coordinates $\varphi(p) = (x_1(p), \cdots, x_n(p))$. If the atlas is maximal in terms of inclusion, then it is a \emph{differentiable structure}. Each atlas for $M$ is included in a maximal atlas. Due to the differentiable structure, the maps in $\mathscr{A}$ are also compatible with all maps of $\mathscr{A}_{\text{max}}$. It is sufficient to describe smooth manifolds by a non maximal atlas.

As long as the activation functions are of the same differentiable structure, they also are compatible with each other, and the manifold does not change as the data propagates through the layers. Only the coordinate system changes.

@@

We apply a theorem from Lie theory, which states that each connected Abelian Lie group $(M,\bullet)$ of dimension $\dim M = n$ is isomorphic to a product space $\mathbb{R}^p \times \mathbb{T}^q$ with $p+q =n$, for a proof we refer to \cite[p.~116]{onishchik1993lie}. The $q$-torus $\mathbb{T}^q$ is a surface of revolution. It moves a curve around an axis of rotation. These axes are given by $1$-spheres, such that the $q$-torus is a product space $\mathbb{T}^q = \mathcal{S}^1_1 \times \cdots \times \mathcal{S}^1_q$. The initial decomposition of connected commutative Lie groups can be further simplified into $M \simeq \mathbb{R}^p \times (\mathcal{S}^1_1 \times \cdots \times \mathcal{S}^1_q)$. Recall, that the $(n-1)$-sphere is given by $\mathcal{S}^{n-1} := \{x \in \mathbb{R}^n \; \vert \; \norm{x}_2 = 1\}$. Thus, $\mathcal{S}^1$ can be embedded into $\mathbb{R}^2$. Next, we derive how many dimensions are at least needed for a suitable embedding. For each $1$-sphere we count two dimensions and for each real line accordingly one. Finally, we have to estimate from data how the decomposition may look like to have the topology of the data manifold.

\subsection{Persistent homology}
Algebraic topology provides a computable tool to study not the topology of a set of points themselfs but an Abelian group attached to them. The core interest in this discipline lies in \emph{homotopy equivalences}, an equivalence class to which objects belong that are continuously deformable into one another. This is much broader than homeomorphism. For two topological spaces $X$ and $Y$ we seek for a function $h: X \times I \rightarrow Y$, which gives the identity at time $h_0(X) = X$ and for $h_1(X) = X$ a mapping into another topological space. If this mapping is continuous with respect to its arguments, it is called \emph{homotopy}. Consider two functions $f,g: I \rightarrow X$ so that $f(1) = g(0)$. Then there is a \emph{composition} of product paths $f \cdot g$, which pass first through $f$ and then through $g$ and which is defined as $f \cdot g (s) = f(2s)$ for $0 \leq s \leq 1/2$ and $g(2s - 1)$, for $1/2 \leq s \leq 1$. We first run $f$ at double speed up to $1/2$ and from $1/2$ to $1$ we run the function $g$ at double speed. In addition, suppose that a family of functions $f: I \rightarrow X$ is given which have the same start and end point $f(0) = f(1) = x_0 \in X$. They intuitively form a \emph{loop}. The set of all homotopy classes $[f]$ of loops $f: I \rightarrow X$ at base point $x_0$ is noted as $\pi_1(X,x_0)$ and is called first homotopy group or fundamental group. The group operation is the product of equivalence classes $[f][g] = [f \cdot g]$. If we do not look at the interval $I$, but at mappings considering the unit cube $I^n$, we obtain the $n$-th homotopy group $\pi_n(X,x_0)$ by analogy. With the help of homotopy groups we study connected components of a topological space for the $0$th group, the loops for the $1$th group, the cavities for the $2$th and so forth. Since homotopy groups are difficult to compute, we resort to an algebraic variant, the \emph{homology groups}.

@@

\input{figure1.tex}
\subsubsection{Simplicial complexes} A \emph{simplicial complex} $K$ is a set of simplices, such that any face of a simplex of $K$ is a simplex of $K$ and the intersection of any two simplices of $K$ is either empty or a common face of both \cite[p.~11]{boissonnat2018geometric}. We require every $(k+1)$-element, with $k < n$ to be a $k$-simplex of $[\sigma] \subset K$. This defines the simplicial complex as a topological space. These properties ensure that simplices are added to the complex in a way, such that they are \emph{glued} edge to edge, point to point, and surface to surface. Thus, every $n$-simplex has $n+1$ distinct vertices and no other $n$-simplex has the same set of vertices. This is a unique combinatorial description of vertices together with a collection of sets $\{[\sigma]_1, \cdots, [\sigma]_m\}$ of $k$-simplices, which are $k+1$ element subsets of $K$ \cite[p.~107]{hatcher2002algebraic}. This set system can be realized geometrically with a homeomorphism into Euclidean space.

\subsubsection{Associated Abelian groups} Algebraically, the simplices can be realized as a system of linear combinations as $\sum_i \lambda_i [\sigma]_{i}$ following Eq. \ref{simplex}. Together with addition as an operation, they form a group structure called $k$th chain group $\langle C_k, +\rangle$ for the corresponding dimension of the $k$-simplices used. The group is commutative or Abelian. The used coefficients $\lambda_i$ induce an orientation, intuitively a direction of the edges, by negative and positive signs, such that $[\sigma] = -[\tau]$, iff $[\sigma] = [\tau]$ and $[\sigma]$ and $[\tau]$ have different orientations. The objects of the chain group are called $k$-chains, for $[\sigma] \in C_k$. The groups are connected by a homomorphism, a mapping which respects the algebraic structure, the boundary operator: $\partial_k: C_k \rightarrow C_{k-1}$, $\partial_k [\sigma] \mapsto \sum_i (-1)^i [v_0,v_1,\cdots,\hat{v}_i, \cdots, v_n]$. The $i$th face is omitted, alternatingly. Using the boundary operator we yield the following sequence of Abelian groups, which is a chain complex cf. Fig. \ref{figure1} a):
\begin{equation}
    0 \xrightarrow[]{\partial_{k+1}} C_k \xrightarrow[]{\partial_{k}} C_{k-1} \xrightarrow[]{\partial_{k-1}} \cdots \xrightarrow[]{\partial_{2}} C_1 \xrightarrow[]{\partial_{1}} C_0 \xrightarrow[]{\partial_{0}} 0.
\end{equation}

@@

\section{Neural nets}
\label{sec:neuralnets}
Neural nets are a composition of \emph{affine transformations} with a non-linear activation function. This transformation obtains collinearity. It also preserves parallelism and partial relationships. The projection onto the $(l+1)$-th layer of such a net can be formalized as $\textbf{x}^{(l+1)} = f(\mathbf{W}^{(l+1)} \cdot \textbf{x}^{(l)}+\textbf{b}^{(l+1)}).$ The composition of multiple such maps is a deep neural net.

@@

\section{Experimental setting}
We train an auto-encoder using as input heavily noisy images and map them to their noiseless original. We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$, such that the input is weighted $0.5 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$. The embedding dimension is increased for each experiment by $2$, i.e. $2,4,6, \cdots, 784$. $392$ neural nets have been trained $\times 100$. Each line of Fig. \ref{figure2} c), d) represents one neural architecture as averaged loss function.
\input{figure2.tex}
\subsubsection*{Persistence landscapes} We use the Delaunay-complex for the filtration, according to Melodia et al. \cite{lume}. The maximum $\alpha$-square is set to $e^{12}$. The maximal expansion dimension of the simplicial complex is set to $10$. The maximal edge length is set to $0.1$. The persistent landscapes are smoothened by a Gaussian filter $G(x) = 1 / \sqrt{2 \pi \sigma^2} \cdot e^{- x^2 / 2 \sigma^2}$ with $\sigma = 2$ for visualization purposes. We implement the persistence diagrams and persistence landscapes using the \texttt{GUDHI} library \cite{gudhi:urm}. Persistent homology is computed on a \texttt{NVIDIA Quadro P4000}.
\subsubsection*{Hyperparameters} The batch size is set to $128$, which should help to better recognize good parametrization. Higher batch size causes explosions of the gradient, which we visualize in Fig. \ref{figure2} c), d). For optimization we use \texttt{ADAM} \cite{KingmaB14} and a learning rate of $10^{-4}$. The dataset is randomly shuffled. We implement in \texttt{Python} $3.7$. We use \texttt{Tensorflow} as backend \cite{tensorflow2015-whitepaper} with \texttt{Keras} as wrapper \cite{chollet2015keras}. The neural training is conducted on \texttt{Intel Core i}$7$ $9700$\texttt{K} processors.
\input{figure3.tex}
\subsubsection*{Neural net} We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = 1/n \sum_{i=1}^{n} (\textbf{x}^{(0)}_i-\textbf{y}_i)^2$ as loss function. The sigmoid activation function $\sigma(\mathbf{x}^{(l)}) = 1 / 1 + e^{-\mathbf{x}^{(l)}}$ is applied throughout the net to yield the structure of a smooth manifold. Dense neural nets are used with bias term. The datasets are flattened from $(32,32,3)$ into $(3072)$. The first layer is applied to the flattened dataset, which is also adapted by back propagation, i.e. $\varphi^{(0)}(x^{(0)}) =$ $\sigma(\sum_{i=1}^{n} \textbf{W}_{ji}^{(1)} \textbf{x}_{i}^{(0)} + \textbf{b}_{i}^{(1)})$. It transfers the data to the desired embedding dimension. The subsequent layers are implemented as residual invertible layers following Dinh et al. \cite{DinhSB17}. We use invertible architectures to stay on the same differentiable structure during learning. A total of $5$ hidden layers are used, all with the same embedding dimension.
